---
title: "Alumni Donation Driver Analysis"
author: "Me"
date: "`r Sys.Date()`"
output: html_document
---

```{r message=FALSE}
library(DALEX)
library(gbm)
library(regclass)
library(tidyverse)
library(randomForest)
library(caret)
library(ggplot2)
library(dplyr)
library(dplyr)
library(ggplot2)
library(inTrees)
library(arules)
library(rpart)
library(rpart.plot)

```

This document provides examples of partial dependence plots, interaction plots, and breakdown plots for the alumni donations case.  There are two datasets:  `CASH` (logarithm of the total amount of cash an alumnus has given so far) and `DONATED` (whether someone donated in 2023).



```{r}
load("BAS479-AlumniFinalCaseF24.RData")
CASH$ID <- NULL
DONATED$ID <- NULL
```


## (Option 1) Understanding predictors for DonatedIn2023

Setup:

```{r}
TRAIN.PREDICTORS <- DONATED
TRAIN.PREDICTORS$DonatedIn2023 <- NULL
TRAIN.TARGET <- as.numeric(DONATED$DonatedIn2023 == "Yes")


write.csv(DONATED, "/Users/husna/Downloads/BAS479/Donated.csv", row.names = FALSE)


```

```{r}
# Example building an explainer with random forest (takes a while)
set.seed(479); FOREST <- randomForest(
  DonatedIn2023 ~ ., 
  data = DONATED, 
  mtry = 30
)

donated_forest_exp <- DALEX::explain(FOREST, data = TRAIN.PREDICTORS, y = TRAIN.TARGET)
shop_explainer_donate <- donated_forest_exp
shop_vi <- model_parts(shop_explainer_donate)


# A lower Gini value means the node is "pure" (all data points belong to the same class), and a higher Gini means the node is impure (data points are split across multiple classes).
# Higher MDG values: A feature with a high Mean Decrease Gini value is considered more important because it contributes more to reducing the uncertainty in predicting the target variable.
# Lower MDG values: A feature with a low MDG value is considered less important since it does not significantly help in reducing the impurity
importance(FOREST)

# Get the feature importance from the random forest model
importance_values <- importance(FOREST)

# Sort the feature importance values in descending order
sorted_importance <- importance_values[order(importance_values[, 1], decreasing = TRUE), ]

# Print the sorted importance
print(sorted_importance)

# Output

# NumYearsDonate, BequestLikelihood, TotalCash, AnnuityLikelihood, LifetimeCashGift, Atheltics Donor, CRTLikelihood, Amount1stGift, LifetimeCashGiftCategory, GradYear





# Step 1: Get the feature importance from the random forest model
importance_values <- importance(FOREST)

# Step 2: Calculate the total importance (sum of all feature importance values)
total_importance <- sum(importance_values[, 1])

# Step 3: Convert the importance values to percentages
importance_percentage <- (importance_values[, 1] / total_importance) * 100

# Step 4: Create a new data frame with features and their percentage importance
importance_df <- data.frame(Feature = rownames(importance_values), 
                            Importance = importance_percentage)

# Step 5: Sort the data frame by percentage importance in descending order
importance_df <- importance_df[order(importance_df$Importance, decreasing = TRUE), ]

# Step 6: Print the sorted importance percentages
print(importance_df)


importance_df_top10 <- importance_df %>%
                        filter(importance_df$Importance > 4.5)

# Create a bar plot to visualize feature importance percentages
feature_plot_top5 <- ggplot(importance_df_top10, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "#5DADE2") +
  coord_flip() +  # Flip axes for better readability
  labs(title = "Feature Importance Breakdown", x = "Feature", y = "Importance (%)") +
  theme_minimal() +
  theme(
        axis.text.x = element_text(angle = 90, hjust = 1, size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 18),
        plot.background = element_rect(fill = "lightgray", color = "black"),  # Change plot background color
        panel.background = element_rect(fill = "white", color = "black"),    # Change panel background color
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),  # Customize grid lines
        panel.grid.minor = element_blank()  # Remove minor grid lines
        
        )

plot(feature_plot_top5)

ggsave("feature_importance_plot_top5.png", plot = feature_plot_top5, width = 14, height = 10, dpi = 300)





```

```{r}
# Creating random forest without 3rd party predictors

TRAIN.PREDICTORS_REDUCED <- TRAIN.PREDICTORS %>%
  select(-EstGiftCapacity, -EstIncome, -EstNetWorth, -EstTotalAssets,
         -AnnuityLikelihood, -BequestLikelihood, -CRTLikelihood)

set.seed(479)
FOREST_REDUCED <- randomForest(
  DonatedIn2023 ~ ., 
  data = cbind(TRAIN.PREDICTORS_REDUCED, DonatedIn2023 = DONATED$DonatedIn2023),
  mtry = 30
)


donated_forest_exp_reduced <- DALEX::explain(FOREST_REDUCED, data = TRAIN.PREDICTORS_REDUCED, y = TRAIN.TARGET)
shop_vi_reduced <- model_parts(donated_forest_exp_reduced)


```


```{r}


TRAIN.PREDICTORS_Donated <- TRAIN.PREDICTORS

TRAIN.PREDICTORS_REDUCED_Donated <- TRAIN.PREDICTORS_REDUCED


TRAIN.PREDICTORS_Donated$DonatedIn2023 <- TRAIN.TARGET

TRAIN.PREDICTORS_REDUCED_Donated$DonatedIn2023 <- TRAIN.TARGET


TRAIN.PREDICTORS_Donated$DonatedIn2023 <- as.factor(TRAIN.PREDICTORS_Donated$DonatedIn2023)
TRAIN.PREDICTORS_REDUCED_Donated$DonatedIn2023 <- as.factor(TRAIN.PREDICTORS_REDUCED_Donated$DonatedIn2023)





full_model_cv <- train(
  DonatedIn2023 ~ ., 
  data = TRAIN.PREDICTORS_Donated, 
  method = "rf", 
  trControl = trainControl(method = "cv", number = 5)
)

# Train full model
full_model_cv <- train(
  DonatedIn2023 ~ ., data = TRAIN.PREDICTORS_Donated,
  method = "rf", trControl = trainControl(method = "cv", number = 5)
)

# Train reduced model
reduced_model_cv <- train(
  DonatedIn2023 ~ ., data = TRAIN.PREDICTORS_REDUCED_Donated,
  method = "rf", trControl = trainControl(method = "cv", number = 5)
)

# Compare metrics
cat("Full Model Accuracy:", max(full_model_cv$results$Accuracy), "\n")
cat("Reduced Model Accuracy:", max(reduced_model_cv$results$Accuracy), "\n")


```


```{r}
# feature importance for reduced

shop_vi_reduced_sorted <- shop_vi_reduced %>% arrange(desc(dropout_loss))




# Filter the data using dplyr
reduced_shop_vi_df <- shop_vi_reduced %>%
filter(dropout_loss > 0.0002)


#reduced_feature <- plot(shop_vi_reduced)

reduced_feature <- ggplot(reduced_shop_vi_df, aes(y = reorder(variable, dropout_loss), x = dropout_loss)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(title = "Feature Importance",
         y = "Features", x = "Importance (dropout_loss)") +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 18),
        plot.background = element_rect(fill = "lightgray", color = "black"),  # Change plot background color
        panel.background = element_rect(fill = "white", color = "black"),    # Change panel background color
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),  # Customize grid lines
        panel.grid.minor = element_blank()  # Remove minor grid lines
    )
ggsave("reduced_feature_importance_plot.png", plot = reduced_feature, width = 14, height = 10, dpi = 300)


```


```{r}

baseline_metrics <- model_performance(shop_explainer_donate)
reduced_metrics <- model_performance(donated_forest_exp_reduced)

print(baseline_metrics)
print(reduced_metrics)

```
```{r}

metric_diff <- baseline_metrics$auc - reduced_metrics$auc
print(paste("Change in AUC:", metric_diff))


```


```{r}
plot(shop_vi)


# low dropout levels not important
print(shop_vi)



# Sort and print in descending order by dropout_loss
shop_vi_sorted <- shop_vi %>% arrange(desc(dropout_loss))
print(shop_vi_sorted)


      
```

```{r}
# For combinations of two features
partial_combos_top2 <- model_profile(
  shop_explainer_donate, 
  variables = c("NumYearsDonate", "Amount1stGift")
)

plot(partial_combos_top5)




# Predict probabilities of donating

TRAIN.PREDICTORS.COPY <- TRAIN.PREDICTORS

TRAIN.PREDICTORS$donation_likelihood <- predict(FOREST, TRAIN.PREDICTORS, type = "prob")[, 2]



# Extract rules from the random forest model
rules <- extractRules(FOREST, TRAIN.PREDICTORS)

# Present the rules in a readable format
readable_rules <- presentRules(rules, TRAIN.PREDICTORS)
head(readable_rules)



```

```{r}



# Train a decision tree
donation_tree <- rpart(DonatedIn2023 ~ EstGiftCapacity + EstIncome + EstNetWorth + 
                       EstTotalAssets + AnnuityLikelihood + BequestLikelihood + CRTLikelihood,
                       data = train_data, method = "class")

# Plot the decision tree
plot(donation_tree)
text(donation_tree, use.n = TRUE)

```


```{r}




train_data <- data.frame(TRAIN.PREDICTORS, DonatedIn2023 = TRAIN.TARGET)

# Build the decision tree model
tree_model <- rpart(DonatedIn2023 ~ ., data = train_data, method = "class")#, control = rpart.control(minsplit = 5, maxdepth = 5))




print(tree_model)

plot(tree_model)

# Visualize the decision tree
rpart.plot(tree_model, main = "Decision Tree for DonatedIn2023")

# Evaluate the model on training data
train_predictions <- predict(tree_model, TRAIN.PREDICTORS, type = "class")
conf_matrix <- table(Predicted = train_predictions, Actual = TRAIN.TARGET)

# Print the confusion matrix

print(conf_matrix)



print(tree_model)

# tree shows class distribution 




# if the num years is less than 2.5 then their chance of donating is 5%
#

# percentage is the likelohood to not donate 


levels(DONATED$AgeAtFirstGift)

mean(DONATED$DonatedIn2023)
```

NumYearsDonate (How many times out of the last 9 years have they donated) is the most important variable to the model, followed by the first gift amount given to UT, graduation year, and being an athletics donor.

```{r}

#plot(shop_vi)
library(dplyr)

library(ggplot2)
p <- plot(shop_vi)
p <- ggplot(shop_vi, aes(y = reorder(variable, dropout_loss), x = dropout_loss)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    labs(title = "Feature Importance",
         y = "Features", x = "Importance (dropout_loss)") +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1, size = 14),
        axis.title = element_text(size = 16),
        plot.title = element_text(size = 18),
        plot.background = element_rect(fill = "lightgray", color = "black"),  # Change plot background color
        panel.background = element_rect(fill = "white", color = "black"),    # Change panel background color
        panel.grid.major = element_line(color = "gray", linetype = "dashed"),  # Customize grid lines
        panel.grid.minor = element_blank()  # Remove minor grid lines
    )
ggsave("feature_importance_plot.png", plot = p, width = 14, height = 10, dpi = 300)
```

The model's predicted probability of donating has a logical steadily increasing relationship with the number of years donated (out of the last 9).  As long as they've donated in at least one of the last 9 years, the probability of donating is above the overall averages (12.68%). 

```{r}
plot(DonatedIn2023 ~ NumYearsDonate, data=DONATED); abline(h=0.1268)
plot(model_profile(shop_explainer_donate, type = "accumulated"), variables = "NumYearsDonate")
```
```{r}
# Create model profile for the "NumYearsDonate" feature
partial_BequestLikelihood <- model_profile(shop_explainer_donate, type = "accumulated", variables = "BequestLikelihood")
plot(partial_numyearsdonate) +
  labs(title = "Accumulated Partial Dependence Plot for Bequest Likelihood",
       x = "BequestLikelihood", y = "Predicted Response") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}
partial_TotalCash <- model_profile(shop_explainer_donate, type = "accumulated", variables = "TotalCash")
plot(partial_TotalCash) +
  labs(title = "Accumulated Partial Dependence Plot for Total Cash Donated",
       x = "Total Cash", y = "Predicted Response") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
partial_AnnuityLikelihood <- model_profile(shop_explainer_donate, type = "accumulated", variables = "AnnuityLikelihood")
plot(partial_AnnuityLikelihood) +
  labs(title = "Accumulated Partial Dependence Plot for Annuity Likelihood",
       x = "AnnuityLikelihood", y = "Predicted Response") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`GradYear` is the year a student received their 1st degree from UT. We see very high chance of donation from recent grads.  This might be attributed to senior gifts.  Otherwise, the probability of donating in 2023 steadily decreases the farther back in time their graduation year was, until about 1986, where the probability of donating in 2023 starts to get higher for earlier graduation years.

```{r}
partial_gradyear <- plot(model_profile(shop_explainer_donate, type = "accumulated"), variables = "GradYear")
plot(partial_gradyear) +
  labs(title = "Accumulated Partial Dependence Plot for GradYear",
       x = "GradYear", y = "Predicted Response") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

`Amount1stGift` is the (log10 of the) amount of an alumni's first gift to UT.  The predicted probability of donating increases steady with this amount.

```{r}
partial_amount1stgift <- plot(model_profile(shop_explainer_donate, type = "accumulated"), variables = "Amount1stGift")
plot(partial_amount1stgift) +
  labs(title = "Accumulated Partial Dependence Plot for Amount1stGift",
       x = "Amount1stGift", y = "Predicted Response") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

There's an interaction going on between `NumYearsDonate` and `School1stDegree` (HBB or Other).  When the number of years previously donated is 0-2, the predicted probability of donating among HBB graduates is higher than the predicted probability for other schools.  However, between 3-6 years, the opposite is true. 

```{r}
set.seed(411); profile_group <- model_profile(
  explainer = shop_explainer_donate,
  variables = c("NumYearsDonate"),
  groups = "School1stDegree", type = "partial"
)
plot(profile_group)
```

`CRTLikelihood.`  See https://kb.blackbaud.com.au/articles/Article/94948 This is information UT pays for.  It's a "score" for various types of donations.

There is no interaction between `Gender` and `CRTLikelihood` according to the model.  While the predicted probability of donating among women is always higher, the bends/curves are basically the same and the difference in predicted probabilities doesn't vary over the CRT score.  

```{r}
set.seed(2024);profile_group <- model_profile(
  explainer = shop_explainer_donate,
  variables = c("CRTLikelihood"),
  groups = "Gender", type = "partial"
)
plot(profile_group)
```


To make a breakdown plot, we first want to find some "interesting" cases.  It would be useful to understand why a particular alumnus has a high predicted probability of donating in 2023 and why a particular alumnus has a low one.


```{r}
set.seed(479); predictions <- predict(FOREST, newdata = DONATED, type = "prob")[, 2]
  #Recall the [,2] is because predict gives us a dataframe of No/Yes predictions, one 
  #column for each class, and the Yes is the 2nd column
high_prob_donate <- as.numeric( which(predictions > 0.963 & predictions < 0.965) )
low_prob_donate <- as.numeric( which(predictions > 0.019 & predictions < 0.021) )
```

In the final case, you'll pick one from each set and interpret the breakdown plot to provide understanding about why those alumni have high/low predicted values.

For illustration here, let's look at how the model arrived at its prediction for a random person (how about row 2684, even though it doesn't correspond to a particularly high or low probability!).

```{r}
#Partial dependence plots take a while for a dataset of this size
specific.alumnus <- DONATED[2684, ]
set.seed(479); plot(predict_parts(explainer = shop_explainer_donate, new_observation = specific.alumnus, type = "break_down"))
```
```{r}

# Identify rows with high and low probabilities
high_pred_row <- which.max(predict(shop_explainer_donate, DONATED))  # Row with highest prediction
low_pred_row <- which.min(predict(shop_explainer_donate, DONATED))  # Row with lowest prediction

# Extract specific observations
high_pred_alumnus <- DONATED[high_pred_row, ]
low_pred_alumnus <- DONATED[low_pred_row, ]

# Generate breakdown plots
set.seed(479)
high_pred_plot <- predict_parts(explainer = shop_explainer_donate, new_observation = high_pred_alumnus, type = "break_down")
low_pred_plot <- predict_parts(explainer = shop_explainer_donate, new_observation = low_pred_alumnus, type = "break_down")

# Plot the breakdowns
plot(high_pred_plot, title = "Breakdown Plot for High Prediction")
plot(low_pred_plot, title = "Breakdown Plot for Low Prediction")


```


Our default prediction is ~13% chance of donation.  

This alumnus just graduated in 2023, which provides a big boost to the predicted probability (up 30%).  A bunch of variables related to the fact that alumnus hasn't donated yet (lifetime cash is 0, lifetimecashgiftcategory none, totalcash=none, when1stgift=none, amount1stgift 0) adjusts the probability up and down by some percentage points as well.  

Since the alumnus didn't participate in any student activities, his predicted donation probability decreases by about 7%.  Being male decreases the predicted donation probability an additional 5% for this alumnus.  When all is said and done, the predicted probability is 18%, a still a little above average.

One really cool thing about breakdown plots is they show how contribution of a variable to predictions can differ a lot when compared to trends in the overall data.  In this example, the alumnus being male decreased the predicted probability by 5%, even though overall donate rate among men (12.9%) is higher than women (12.2%).
```{r}



library(pROC)
roc_curve <- roc(TRAIN.TARGET, train_predictions)
auc(roc_curve)

```


```{r}
aggregate(DonatedIn2023=="Yes" ~ Gender, data=DONATED,FUN=mean)



# Breakdown of Top Drivers


aggregate(DonatedIn2023=="Yes" ~ NumYearsDonate, data=DONATED,FUN=mean)

aggregate(DonatedIn2023=="Yes" ~ Amount1stGift, data=DONATED,FUN=mean)

aggregate(DonatedIn2023=="Yes" ~ AthleticsDonor, data=DONATED,FUN=mean)

aggregate(DonatedIn2023=="Yes" ~ GradYear, data=DONATED,FUN=mean)

aggregate(DonatedIn2023=="Yes" ~ LifeTimeCashGiftCategory, data=DONATED,FUN=mean)



```

```{r}
TREE <- rpart( ~ ., data = BANK, minbucket = 30, cp =  0.020618557)
TREE$cptable
TREE
visualize_model(TREE)
```

```{r}


# Prepare the data
# Convert the dataset into transactions format
# Here, all features should be categorical for association rules
donated_trans <- as(as.matrix(DONATED), "transactions")


rules <- apriori(
  DONATED, 
  parameter = list(support = 0.01, confidence = 0.6)  # Adjust thresholds as needed
)



# Inspect top rules
inspect(sort(rules, by = "lift")[1:10])  # Top 10 rules by lift


unique(rhs(rules))

# Filter rules targeting 'DonatedIn2023 = Yes'
target_rules <- subset(rules, rhs %in% c("DonatedIn2023=1"))
inspect(sort(target_rules, by = "confidence")[1:10])

 # Top 10 rules by confidence


# Check current levels of DonatedIn2023
levels(DONATED$DonatedIn2023)

# Relevel DonatedIn2023 to Yes and No
levels(DONATED$DonatedIn2023) <- c("Yes", "No")

# Verify the change
levels(DONATED$DonatedIn2023)


target_rules <- subset(rules, rhs %in% c("DonatedIn2023=Yes"))  # Adjust based on your unique rhs labels

# Sort and inspect the top 10 rules by confidence
inspect(sort(target_rules, by = "confidence")[1:10])


```













## (Option 2) Understanding predictors for `LogTotalCash`

Setup the data:

```{r}
TRAIN.PREDICTORS <- CASH
TRAIN.PREDICTORS$LogTotalCash <- NULL
TRAIN.TARGET <- CASH$LogTotalCash
```

Build the explainer and check out variable importances.

```{r}
# Example building an explainer with random forest
set.seed(479); FOREST <- randomForest(LogTotalCash ~ ., data = CASH )
shop_explainer <- explain(FOREST, data = TRAIN.PREDICTORS, y = TRAIN.TARGET)
shop_vi <- model_parts(shop_explainer, loss_function = loss_root_mean_square)
```

`DonatedInLast10Years` is the most important variable for prediction in the random forest model.  `DonatedInLast10Years` definition is "did they give a cash gift at any point in the last 10 years?", so this does make sense.  This quantity is very important for predicting the lifetime amount that someone has donated, and the variable importance plot lets us know the models performance suffers drastically if it is omitted.

```{r}
plot(shop_vi)
```

`BequestLikelihood` (see https://kb.blackbaud.com.au/articles/Article/94948 for more detail) is a piece of information pays of a 3rd party for.   It's a score that reveals how likely the company things the alumnus would leave a gift to UT as part of a will or a trust.

When this score is nearing 1000 we see the model predicting about 2.5 for `LogTotalCash` on average.  This translates to a prediction of 320 dollars (10^2.5).  For scores near 300, the model predicts about 2.15 for `LogTotalCash`, translating into a prediction of about 140 dollars.  

```{r}
set.seed(479); plot(
  model_profile(shop_explainer, 
                type = "partial"), 
  variables = "BequestLikelihood"
)
10^2.5
10^2.15
```

Looking at `BequestLikelihood` and `DonatedInLast10Years` together we see that `DonatedInLast10Years` gives a boost to predictions as expected.  If an individual has `DonatedInLast10Years` and has a `BequestLikelihood` close to 1000, we expect to see a log donation size of 2.7. Which translates to donation size of $500.

This boost looks go *grow* in size as bequest likelihood score increases (indicating there's a bit of an interaction between the variables).  At low bequest scores (near 500), the predictions are about 2.2 vs. 2.1 (158 vs 126 dollars).  At high bequest (near 1000), the predictions are about 2.7 vs. 2.25 (500 vs 178).  Thus, donating in the last 10 years has more of an effect on the predicted value when the bequest likelihood is high than small.

```{r}
set.seed(479); profile_group <- model_profile(
  explainer = shop_explainer,
  variables = c("BequestLikelihood"),
  groups = "DonatedInLast10Years", 
  type = "partial"
)
plot(profile_group)
```

Another predictor with an interaction with bequest likelihood is `FavoriteSport`.  For alumni whose favorite game is women's basketball, the bequest likelihood doesn't affect predictions (flat at around 3, or 1000).  For alumni whose favorite game is women's basketball, the predictions sharply increase at a bequest score of 750 (intriguing), while the predictions steadily increase with bequest score when favorite game is "None"  or "Football".

```{r}
set.seed(2023); profile_group <- model_profile(
  explainer = shop_explainer,
  variables = c("BequestLikelihood"),
  groups = "FavoriteSport", type = "partial"
)
plot(profile_group)
```

`EstGiftCapacity`:  information purchased from a different third party that provides an estimate of the total amount of money the alumnus could give over the course of his/her lifetime.  Many levels are very rare.

Here's where the DALEX package is a bit of a letdown since so many colors are so similar.  Estimated gift capacity seems to matter a lot more to predictions at low values of bequest score (big spread in predictions based on estimated wealth capacity) and less of a difference at large bequest scores (lines are more scrunched together). 

```{r}
set.seed(2023); profile_group <- model_profile(
  explainer = shop_explainer,
  variables = c("BequestLikelihood"),
  groups = "EstGiftCapacity", type = "partial"
)
plot(profile_group)
```

OpenReceiveRatio - this records the ratio of the number of emails sent by the development office that were opened by the alumnus to the number of emails that were received.  For example, if the alumus received 50 emails and opened 40 of them, the ratio would be 0.8.  Since it's possible to open an email multiple times, sometimes this ratio is above 1.  Most values are between 0-1.5 in the data.

Surprisingly, for people that have donated in the last 10 years, the model predicts less total cash donated as the open to receive ratio increases (beyond about 0.6).  This doesn't make much intuitive sense (we'd expect total cash donated to increase with this ratio as it does for the "hasn't donated in last decade" class).  

```{r}
hist(CASH$OpenReceiveRatio)
set.seed(2023);  profile_group <- model_profile(
  explainer = shop_explainer,
  variables = c("OpenReceiveRatio"),
  groups = "DonatedInLast10Years", type = "partial"
)
plot(profile_group)
```


To make a breakdown plot, we first want to find some "interesting" cases.  It would be useful to understand why a particular alumnus has a high predicted total cash value and why a particular alumnus has a low one.

```{r}
set.seed(479); predictions <- predict(FOREST, newdata = CASH)
# Find people that the model thinks will donate a lot of money
expected_high_donation <- as.numeric( which(predictions > 3.8) )

# Find people that the model thinks will donate a small amount of money
expected_small_donation <- as.numeric( which(predictions < 0.9) )
```

In the final case, you'll pick one from each set and interpret the breakdown plot to provide understanding about why those alumni have high/low predicted values.

For illustration here, let's look at how the model arrived at its prediction for a random person (how about row 2022, though it's not a particularly interesting case!).


```{r}
specific.alumnus <- CASH[2022, ]
plot(predict_parts(explainer = shop_explainer, new_observation = specific.alumnus, type = "break_down"))
```


Our default prediction is 2.307 ($200).  They're a young alumnus who just graduated (born 1999 and graduated 2022), which decrease their prediction down by 0.239 and by 0.246, respectively, or down to 10^(2.307-0.239) = 117 dollars then down to 10^(2.307-0.239-0.246) = 66 dollars.  

Other factors such as marital status (unknown), estimated gift capacity (unable to rate), small bequest likelihood, no student activities, etc., continue to drag down the final prediction down to 1.236, or 10^(1.236) = 17.22.


Let's try another (rather uninteresting) alumnus:

```{r}
specific.alumnus <- CASH[1337, ]
plot(predict_parts(explainer = shop_explainer, new_observation = specific.alumnus, type = "break_down"))
```

While this alumnus is clearly very well-off (estimated gift capacity of 100-200K, net worth of 1-5 million), since they haven't donated in the last 10 years and because they have relatively low bequest and annuity scores, have no family relations going to UT, and have relatively low values for the email stats, their final prediction ends up at 1.735, or 54.

